---
title: An R Markdown document converted from "./vignettes/DPT.ipynb"
output: rmarkdown::html_vignette
bibliography: bibliography.bib
---

```{r}
set.seed(1)
```

Diffusion Pseudo Time (DPT) is a pseudo time metric based on the transition probability of a diffusion process [@haghverdi_diffusion_2016].

*destiny*  supports `DPT` in addition to its primary function of creating `DiffusionMap`s from data.

```{r}
library(destiny)    # load destiny…
data(guo)           # …and sample data
library(gridExtra)  # Also we need grid.arrange
```

`DPT` is in practice independent of Diffusion Maps:

```{r}
par(mar = rep(0, 4))
graph <- igraph::graph_from_literal(
    data -+ 'transition probabilities' -+ DiffusionMap,
    'transition probabilities' -+ DPT)
plot(
    graph, layout = igraph::layout_as_tree,
    vertex.size = 50,
    vertex.color = 'transparent',
    vertex.frame.color = 'transparent',
    vertex.label.color = 'black')
```

However in order not to overcomplicate things, in *destiny*, you have to create `DPT` objects from `DiffusionMap` objects.  
<small>(If you really only need the DPT, skip Diffusion Component creation by specifying `n_eigs = 0`)</small>

```{r}
dm <- DiffusionMap(guo)
dpt <- DPT(dm)
```

The resulting object of a call like this will have three automatically chosen tip cells. You can also specify tip cells:

```{r}
set.seed(4)
dpt_random <- DPT(dm, tips = sample(ncol(guo), 3L))
```

Plotting without parameters results in the DPT of the first root cell:

TODO: wide plot

```{r}
grid.arrange(plot(dpt), plot(dpt_random), ncol = 2)
```

Other possibilities include the DPT from the other tips or everything supported by `plot.DiffusionMap`:

TODO: wide plot

```{r}
grid.arrange(
    plot(dpt, col_by = 'DPT3'),
    plot(dpt, col_by = 'Gata4', pal = viridis::magma),
    ncol = 2
)
```

The `DPT` object also contains a clustering based on the tip cells and DPT, and you can specify where to draw paths from and to:

```{r}
plot(dpt, root = 2, paths_to = c(1,3), col_by = 'branch')
```

You can further divide branches. First simply plot branch colors like we did above, then identify the number of the branch you intend to plot, and then specify it in a subsequent `plot` call. In order to see the new branches best, we specify a `dcs` argument that visually spreads out out all four branches.

```{r}
plot(dpt, col_by = 'branch', divide = 3, dcs = c(-1,-3,2), pch = 20)
```

